---
title: "00-Network-Maps"
output: html_document
date: "2025-05-29"
params:
  city: "Waterloo"
---

This is a parametrized notebook for obtaining the network of the inputted municipality. The components, which are clusters of connected bike lanes (of various forms), are then obtained and can be mapped. To input a city/location, view the `data_sources` file in `\data-raw` within this repository, and input one of the 76 locations from the `municipality` column into the city category under "params" above.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(readr)
library(sf)
library(sfnetworks)
library(ggplot2)
library(igraph)
library(dplyr)
library(r5r)
```

```{r load files}
load(file = glue::glue(here::here(),
                       "/data-raw/classification_dictionary.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/column_descriptions.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/data_sources.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/municipality_infra_summary.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/cycle_network.rda"))
```

```{r select city from parameter}
network <- cycle_network |>
  filter(municipality == {params$city})
```

```{r map cycling network for location}
ggplot(network) +
  geom_sf(geom = network$geom) +
  theme_bw()
```

```{r, eval=FALSE}
unique(cycle_network$source_class)
```

```{r, eval=FALSE}
unique(network$source_class)
# used to just compare what source classes exist for the given city versus the whole dataset
```

```{r converts all network lines into strings}
network_lines <- network |>
  st_cast("LINESTRING", do_split = TRUE)
```

```{r convert network lines into an sf network}
network_lines <- as_sfnetwork(network_lines, directed = FALSE)
```

```{r obtain components from network lines}
components <- components(network_lines)
# https://igraph.org/r/html/1.3.5/components.html
```

```{r check component distribution}
component_distribution(network_lines, cumulative = TRUE)
# not sure about whether to keep cumulative true or false
```

```{r check number of components}
components$no
```

```{r assign IDs to components}
network_lines <- network_lines |>
  activate("nodes") |>
  mutate(component = components$membership)
```

```{r generate nodes and edges of network}
nodes <- st_as_sf(network_lines, "nodes")
edges <- st_as_sf(network_lines, "edges")
```

```{r add node IDs and join to edges}
nodes <- nodes |>
  mutate(node_id = row_number())

edges <- edges |>
  left_join(nodes |>
      st_drop_geometry() |>
      select(node_id, component),
    by = c("from" = "node_id"))
```

```{r map components of network}
ggplot() +
  geom_sf(data = edges,
          aes(color = as.factor(component)), size = 1) +
  scale_color_viridis_d(option = "turbo", guide = "none") +
  theme_bw()
# all colours on https://ggplot2.tidyverse.org/reference/scale_viridis.html
# changing the size value does not appear to change anything
# make sure to add "guide = "none"" or it generates a list of each component's colour that will block the entire map
```
