---
title: "00-Network-Maps"
output: html_document
date: "2025-05-29"
params:
  city: "Waterloo"
---

This is a parametrized notebook for obtaining the network of the inputted municipality. The components, which are clusters of connected bike lanes (of various forms), are then obtained and can be mapped. To input a city/location, view the `data_sources` file in `\data-raw` within this repository, and input one of the 76 locations from the `municipality` column into the city category under "params" above.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(readr)
library(sf)
library(sfnetworks)
library(ggplot2)
library(igraph)
library(dplyr)
library(r5r)
```

```{r load files}
load(file = glue::glue(here::here(),
                       "/data-raw/classification_dictionary.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/column_descriptions.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/data_sources.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/municipality_infra_summary.rda"))

load(file = glue::glue(here::here(),
                       "/data-raw/cycle_network.rda"))
```

```{r select city from parameter}
network <- cycle_network |>
  filter(municipality == {params$city})
```

```{r map cycling network for location}
ggplot(network) +
  geom_sf(geom = network$geom) +
  theme_bw()
```

```{r, eval=FALSE}
unique(cycle_network$source_class)
```

```{r, eval=FALSE}
unique(network$source_class)
# used to just compare what source classes exist for the given city versus the whole dataset
```

```{r set up network}
network <- network |>
  st_union() |>
  st_make_valid() |>
  st_buffer(dist = 30) |> # still trying to determine what buffer distance to use
  st_cast("MULTILINESTRING") |>
  st_line_merge() |>
  st_as_sf()
```

```{r converts all network lines into strings}
network_lines <- network |>
  st_cast("LINESTRING", do_split = TRUE)
# normally do_split is set to TRUE
```

```{r convert network lines into an sf network}
network_lines <- as_sfnetwork(network_lines, directed = FALSE)
# normally directed is set to FALSE
```

```{r, eval=FALSE}
network_lines <- edge.curved=FALSE
# edge.curved normally set to FAlSE
# determine if needed later
```

```{r obtain components from network lines}
components <- components(network_lines)
# https://igraph.org/r/html/1.3.5/components.html
# setting to "strong" or "weak" results in the same number of components
```

```{r check number of components}
components$no
# changes based on the buffer distance
```

```{r check component distribution}
component_distribution <- component_distribution(network_lines, cumulative = TRUE)
component_distribution
# not sure about whether to keep cumulative true or false
# does not work after setting a buffer distance for network
```

```{r plot component distribution}
plot(component_distribution)
```


```{r assign IDs to components}
network_lines <- network_lines |>
  activate("nodes") |>
  mutate(component = components$membership)
```

```{r generate nodes and edges of network}
nodes <- st_as_sf(network_lines, "nodes")
edges <- st_as_sf(network_lines, "edges")
```

```{r add node IDs and join to edges}
nodes <- nodes |>
  mutate(node_id = row_number())

edges <- edges |>
  left_join(nodes |>
      st_drop_geometry() |>
      select(node_id, component),
    by = c("from" = "node_id"))
```

```{r map components of network}
ggplot() +
  geom_sf(data = edges,
          aes(color = as.factor(component)), size = 1) +
  scale_color_viridis_d(option = "turbo", guide = "none") +
  theme_bw()
# all colours on https://ggplot2.tidyverse.org/reference/scale_viridis.html
# changing the size value does not appear to change anything
# make sure to add "guide = "none"" or it generates a list of each component's colour that will block the entire map
# can change "data = edges" to "data = nodes" to show the node central points themselves
# https://ggplot2.tidyverse.org/reference/ggtheme.html
```
